Predicting Conflict Escalation

### 2023-04-28 docs/hes_final/article.qmd
Add future directions

>>>>

We think there are many ways to extend this work. We can group them to align with our initial research and
practical objectives.

### Future research directions

One of the potential research extensions of this work is to continue experimenting
with data sampling and aggregation approaches. We used either monthly averages 
to create $512x1$ vectors or article level samples to create $512x50$ matrices
in our dataset preparation. While we received some promising results, it is possible 
that the first approach makes the data too coarse and the second too noisy. Another
alternative could be generating daily averages and combining them into $512x31$ (largest number of days in a month)
matrices. This could potentially be more sensitive to the signal than the monthly averages
and less noisy than article level samples.

Another direction is dimensionality reduction technique. We observed a clear separation of clusters using a
UMAP technique supervised by the Spike target variable. Using the dataset reduced with this technique
could be a better alternative to a principal component analysis our statistical model uses.

On the transormers side, the output of the encoder also represented a reduced dimension dataset
further fed into the second transformer, the classifier. Using the output of the encoder as input to other
classifiers could be another direction of research.

Finally, due to time and resource constraints in this project we used 20% subset
of daily news articles. Increasing that percentage could potentially lead to better results.

### Future practical directions

Building an application with a user interface could be a natural extension of this work.
We envision a dashboard where a user can view a world map where countries are heatmap colored
according to the estimated probability of escalating conflict. Countries with highest risk
can be flagged or shown to a user in a separate view. In addition to the visual part
of this application it is important to ensure that the model is deployed and
can receive new data input and make inference.

Building a more robust data processing pipeline is another important direction. This would involve
designing either a batch or streaming system that can reliably collect data, preprocess
and aggregate it into required format, train the model and make predictions. The system should handle 
computational resources and cost constraints.
